---
title: "prep-timeseries"
author: "mabarbour"
date: "2021-05-19"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
# set knitr options
knitr::opts_chunk$set(autodep = TRUE, message = FALSE)

# load required libraries
library(MASS) # must be before dplyr
library(tidyverse)
library(cowplot)
library(knitr)

# set ggplot  theme
theme_set(theme_cowplot())
```

# Upload original data

```{r}
original_df <- read_csv("data/InsectAbundanceSurvival.csv")
```

# Underneath leaves correction for LYER and Ptoid

We are only using data from the last week of the experiment, for which we have estimates of LYER and Ptoids (mummies actually) underneath the leaf

```{r}
# only retain cages where LYER or Parasitoids survived
# no surviving BRBR, so we can't estimate this
LYER_df_17 <- filter(original_df, Week == "17") %>% filter(LYER_Survival == 1) %>% dplyr::select(Cage, LYER)
Mummy_df_17 <- filter(original_df, Week == "17") %>% filter(Mummy_Ptoids_Survival == 1) %>% dplyr::select(Cage, Mummy_raw)

# underneath leaf counts, only available for last week of experiment
underneath_df <- read_csv("data/insect_abundance_data_2018-09-26_underneathleaf.csv") # counted to individual resolution

# merge with original data
LYER_df_17 <- left_join(LYER_df_17, dplyr::select(underneath_df, Cage, underneath_LYER = LYER))
Mummy_df_17 <- left_join(Mummy_df_17, dplyr::select(underneath_df, Cage, underneath_Mummy = Mummy))
```

## LYER

```{r }
## exploratory plots ----

# all data
x11(); ggplot(LYER_df_17, aes(x = log1p(LYER), y = log(underneath_LYER))) +
  geom_point() + geom_smooth(method = "lm") 

# remove outlier
x11(); ggplot(LYER_df_17[-31,], aes(x = log1p(LYER), y = log(underneath_LYER))) +
  geom_point() + geom_smooth(method = "lm")

## GLM ----

# I dropped the outlier because I truly think it was an unusual data point and would bias our other estimates.
plot(MASS::glm.nb(underneath_LYER ~ log1p(LYER), LYER_df_17), which = 5)
summary(glm(underneath_LYER ~ log1p(LYER), LYER_df_17[-31,], family = poisson)) # residual deviance suggests a lot of overdispersion, so I'm going to use a negative binomial model
LYER_underneath_glm.nb <- MASS::glm.nb(underneath_LYER ~ log1p(LYER), LYER_df_17[-31,]) 
summary(LYER_underneath_glm.nb)

LYER_correction_glm.nb <- function(x) round(predict(LYER_underneath_glm.nb, newdata = data.frame(LYER = x), type = "response"), 0)
# will add between 30 and 286 individual LYER to LYER_t
LYER_correction_glm.nb(c(0,max(original_df$LYER, na.rm = T)))
```

## Mummy and Ptoids

```{r }
## exploratory plot ----
x11(); ggplot(Mummy_df_17, aes(x = log1p(Mummy_raw), y = log(underneath_Mummy))) +
  geom_point() + geom_smooth(method = "lm")
# the predicted increase is based on 1 data point
summary(Mummy_outlier_glm.nb <- MASS::glm.nb(underneath_Mummy ~ log1p(Mummy_raw), Mummy_df_17))
# hmmm, this model predicts that there could be as many as 43,291 mummies underneath the leaves, which is insane
round(predict(Mummy_outlier_glm.nb, newdata = data.frame(Mummy_raw = c(0, max(original_df$Mummy_raw, na.rm = T))), type = "response"),0)
# I think the baseline value, for which we have 5 data points is reasonable
# but for the increase, I'm going to use the slope of LYER, which seems reasonable given that this was the primary
# consumer-resource relationship and LYER was often found underneath the leaves compared to BRBR

summary(Mummy_underneath_glm.nb <- MASS::glm.nb(underneath_Mummy ~ 1, filter(Mummy_df_17, Mummy_raw == 0)))
round(exp(coef(Mummy_underneath_glm.nb)["(Intercept)"]),0) # predicts 5 underneath leaves if there was a zero above rosette count

Mummy_correction_glm.nb <- function(x) round(exp(coef(Mummy_underneath_glm.nb)["(Intercept)"] + coef(LYER_underneath_glm.nb)["log1p(LYER)"]*log1p(x)),0)
Mummy_correction_glm.nb(c(0,max(original_df$Mummy_raw, na.rm = T))) # now that's more reasonable
```

# Measurement error

Below, I estimate measurement error for BRBR, LYER, and Ptoid counts. This uses data for which we had overlapping counts for MAB (first author) and DTV (technician that was instrumental to the study).

```{r}
# identify cage-week combinations with repeated measurements
rep_df <- original_df %>% 
  distinct(Cage, Week, Counter) %>%
  group_by(Cage, Week) %>%
  summarise(n = n()) %>%
  filter(n > 1)
```

## BRBR

```{r }
# make BRBR measurement error (me) df
BRBR_me_df <- left_join(rep_df, dplyr::select(original_df, Cage, Week, Counter, BRBR, BRBR_Survival)) %>% 
  drop_na() %>% 
  filter(BRBR_Survival == 1) %>%
  spread(key = Counter, value = BRBR) %>%
  mutate(log1p_DTV = log1p(DTV),
         log1p_MAB = log1p(MAB)) %>%
  ungroup()
BRBR_me <- apply(select(BRBR_me_df, log1p_DTV, log1p_MAB), MARGIN = 1, FUN = sd)
hist(BRBR_me)
median(BRBR_me, na.rm = T)
```

## LYER

```{r }
# make LYER measurement error (me) df
LYER_me_df <- left_join(rep_df, dplyr::select(original_df, Cage, Week, Counter, LYER, LYER_Survival)) %>% 
  drop_na() %>% 
  filter(LYER_Survival == 1) %>%
  spread(key = Counter, value = LYER) %>%
  mutate(log1p_DTV = log1p(DTV),
         log1p_MAB = log1p(MAB),
         abs.diff_log1p = log1p_DTV - log1p_MAB,
         mean.diff_log1p = log1p_DTV - (log1p_DTV+log1p_MAB)/2) %>%
  ungroup()
LYER_me <- apply(select(LYER_me_df, log1p_DTV, log1p_MAB), MARGIN = 1, FUN = sd)
hist(LYER_me)
median(LYER_me, na.rm = T)
```

## Mummy-Ptoids

```{r}
# make Mummy_Ptoids measurement error (me) df
Mummy_Ptoids_me_df <- left_join(rep_df, dplyr::select(original_df, Cage, Week, Counter, Mummy_Ptoids, Mummy_Ptoids_Survival)) %>% 
  drop_na() %>% 
  filter(Mummy_Ptoids_Survival == 1) %>%
  spread(key = Counter, value = Mummy_Ptoids) %>%
  mutate(log1p_DTV = log1p(DTV),
         log1p_MAB = log1p(MAB)) %>%
  ungroup()
Mummy_Ptoids_me <- apply(select(Mummy_Ptoids_me_df, log1p_DTV, log1p_MAB), MARGIN = 1, FUN = sd)
hist(Mummy_Ptoids_me)
median(Mummy_Ptoids_me, na.rm = T)
```

# Tidy timeseries data

```{r}
# useful function
mean_integer <- function(x) as.integer(mean(x, na.rm=T)) # remove NA, otherwise, the count will be set to zero if only 1 counter counted.
```

```{r}
# make timeseries dataset
timeseries_df <- original_df %>%
  select(Cage, Week, temp = Temperature, rich = Richness, com = Composition, Col, gsm1, AOP2, AOP2.gsoh, BRBR_Survival, LYER_Survival, Mummy_Ptoids_Survival, BRBR_t = BRBR, LYER_t = LYER, Ptoid_t = Mummy_Ptoids, Mummy_t_raw = Mummy_raw) %>%
  # applying adjustments before averaging across counters
  mutate(LYER_t_adj = ifelse(LYER_Survival == 0 | is.na(LYER_Survival) == T, 1,   # replace extinction zeros with value of 1
                             ifelse(LYER_Survival == 1, LYER_t + LYER_correction_glm.nb(LYER_t), LYER_t)), # add adjusted values
         # Ptoid should always be zero for Week 1 and 2, as they were added after sampling aphids in week 2. But we set them to 1 for extrapolating predictions
         Ptoid_t_adj = ifelse(Mummy_Ptoids_Survival == 0 | is.na(Mummy_Ptoids_Survival) == T, 1,   # replace extinction zeros with value of 1
                              ifelse(Mummy_Ptoids_Survival == 1, Ptoid_t + Mummy_correction_glm.nb(Mummy_t_raw), Ptoid_t))) %>% # add adjusted values
  # apply adjustments for log-log analysis, using minimum count resolutions
  # note that we NEVER used extinct species in fitting the multivariate model, only for extrapolating predictions from multivariate model to extant species
  mutate(BRBR_t = ifelse(BRBR_Survival == 0 | is.na(BRBR_Survival) == T, 1,  # replace extinction zeros with value of 1
                         ifelse(BRBR_t == 0 & BRBR_Survival == 1 & Week < 2, 1,   # replaced undetected zeros before week 2 with value of 1
                                ifelse(BRBR_t == 0 & BRBR_Survival == 1 & Week > 1, 5, BRBR_t))), # replace undetected zeros with value of 5
         LYER_t = ifelse(LYER_Survival == 0 | is.na(LYER_Survival) == T, 1,   # replace extinction zeros with value of 1
                         ifelse(LYER_t == 0 & LYER_Survival == 1 & Week < 2, 1,     # replaced undetected zeros before week 2 with value of 1
                                ifelse(LYER_t == 0 & LYER_Survival == 1 & Week > 1, 5, LYER_t))), # replaced undetected zeros after week 1 with value of 5
         Ptoid_t = ifelse(Mummy_Ptoids_Survival == 0 | is.na(Mummy_Ptoids_Survival) == T, 1,   # replace extinction zeros with value of 1
                          ifelse(Ptoid_t == 0 & Mummy_Ptoids_Survival == 1, 1, Ptoid_t))) %>%     # replaced undetected zeros with value of 1
  group_by(Cage, Week, temp, rich, com, Col, gsm1, AOP2, AOP2.gsoh) %>%
  # average across counters
  summarise_all(~ mean_integer(.)) %>% # funs(mean_integer)
  ungroup()

# initial additions of aphids (4 individuals)
init_timeseries_df <- timeseries_df %>%
  distinct(Cage, temp, rich, com, Col, gsm1, AOP2, AOP2.gsoh) %>%
  # set Ptoid values to 1 for extrapolating predictions
  mutate(Week = 0, BRBR_Survival = 1, LYER_Survival = 1, Mummy_Ptoids_Survival = NA, 
         BRBR_t = 4, LYER_t = 4, Ptoid_t = 1, Mummy_t_raw = NA, # set Mummy_t to be consistent with its raw classification in the data
         LYER_t_adj = 4, Ptoid_t_adj = 1)
init_timeseries_df 

# create lagged (t-1) and leading variable at time t+1
timeseries_df <- bind_rows(init_timeseries_df, timeseries_df) %>%
  arrange(Cage, Week) %>%
  group_by(Cage) %>%
  # leading
  mutate(BRBR_Survival_t1 = lead(BRBR_Survival, n = 1, order_by = Week),
         BRBR_t1 = lead(BRBR_t, n = 1, order_by = Week),
         LYER_Survival_t1 = lead(LYER_Survival, n = 1, order_by = Week),
         LYER_t1 = lead(LYER_t, n = 1, order_by = Week),
         LYER_t1_adj = lead(LYER_t_adj, n = 1, order_by = Week),
         Mummy_Ptoids_Survival_t1 = lead(Mummy_Ptoids_Survival, n = 1, order_by = Week),
         Ptoid_t1 = lead(Ptoid_t, n = 1, order_by = Week),
         Ptoid_t1_adj = lead(Ptoid_t_adj, n = 1, order_by = Week)) %>%
  # lagged. set default = 1 to include in extrapolations for log analysis
  mutate(BRBR_t0 = lag(BRBR_t, n = 1, order_by = Week, default = 1),
         LYER_t0 = lag(LYER_t, n = 1, order_by = Week, default = 1),
         #LYER_t0_adj = lag(LYER_t_adj, n = 1, order_by = Week),
         Ptoid_t0 = lag(Ptoid_t, n = 1, order_by = Week, default = 1))#,
         #Ptoid_t0_adj = lag(Ptoid_t_adj, n = 1, order_by = Week))

# add two parasitoid females to experiment after counts on week 2 and 3
timeseries_df$Ptoid_t[which(timeseries_df$Week == 2)] <- 2
timeseries_df$Ptoid_t[which(timeseries_df$Week == 3)] <- timeseries_df$Ptoid_t[which(timeseries_df$Week == 3)] + 2
# repeat for adjusted
timeseries_df$Ptoid_t_adj[which(timeseries_df$Week == 2)] <- 2
timeseries_df$Ptoid_t_adj[which(timeseries_df$Week == 3)] <- timeseries_df$Ptoid_t_adj[which(timeseries_df$Week == 3)] + 2
# and for lagged values
timeseries_df$Ptoid_t0[which(timeseries_df$Week == 2)] <- 1 # actual zero, but setting to 1 for log() analysis later.
timeseries_df$Ptoid_t0[which(timeseries_df$Week == 3)] <- 2 # 2 parasitoids added in previous week
timeseries_df$Ptoid_t0[which(timeseries_df$Week == 4)] <- timeseries_df$Ptoid_t0[which(timeseries_df$Week == 4)] + 2 # adding 2 parasitoids to count from previous week (often zero)
# and for lagged adjusted values
#timeseries_df$Ptoid_t0_adj[which(timeseries_df$Week == 2)] <- 0 # maintaing at zero, because it is, will adjust by adding 1 for log() analysis later
#timeseries_df$Ptoid_t0_adj[which(timeseries_df$Week == 3)] <- 2 # 2 parasitoids added in previous week
#timeseries_df$Ptoid_t0_adj[which(timeseries_df$Week == 4)] <- timeseries_df$Ptoid_t0_adj[which(timeseries_df$Week == 4)] + 2 # adding 2 parasitoids to count from previous week (often zero)
# before we were ignoring the initial increase of parasitoids from week 2 to week 3, since they are at zeros in the dataset, although I added 2 individuals after week 2 and 3 counts. I added the lines above to incorporate this info
# note that Ptoid_t1 will not exactly match Ptoid_t, but that is intentional since those data are derived from counts

# adjust Mummy_Ptoid_Survival so it doesn't start at Week 3, but Week 2
timeseries_df$Mummy_Ptoids_Survival[which(timeseries_df$Week == 2)] <- 1

# all NA for prediction next week at week 17, which makes sense because that's when the experiment ended so we are going to remove this week
filter(timeseries_df, Week == 17) %>%
  select(BRBR_t1, LYER_t1, LYER_t1_adj, Ptoid_t1, Ptoid_t1_adj) %>%
  data.frame()
timeseries_df <- filter(timeseries_df, Week != 17)

# 17 weeks of data still because we include initial Week = 0
dim(timeseries_df)[1] / length(unique(timeseries_df$Cage)) 

# add new predictors
timeseries_df <- timeseries_df %>%
  mutate(temp = ifelse(temp == "20 C", 0, 3), # now on scale of 1 deg C
         aop2_vs_AOP2 = Col + gsm1 - AOP2 - AOP2.gsoh,
         aop2_genotypes = Col + gsm1,
         AOP2_genotypes = AOP2 + AOP2.gsoh) 
```

# Plant biomass data

```{r}
# load and manage plant biomass time-series data
tmp.biomass <- read_csv("data/ExperimentPlantBiomass.csv") %>%
  select(Cage, Week, No_Plants, Biomass_g) # No_Plants not necessary because we're using Biomass_g as a covariate rather than a response
  # note that there is no biomass data on first week, because plants were left in for 2 weeks
  # we are going to interpolate this biomass based on the plants growth at
  # week 2 of the experiment

# get biomass at t+1
get_biomass_t1 <- tmp.biomass %>%
  mutate(Week = Week - 1) %>%
  select(Cage, Week, Biomass_g_t1 = Biomass_g)

# merge biomass at time t and t+1
merge.biomass.df <- left_join(get_biomass_t1, tmp.biomass) %>%
  # assuming a linear relationship in plant growth between weeks, since
  # they had not hit the saturation point on their growth curve
  mutate(Biomass_g = ifelse(Week == 1, Biomass_g_t1 - Biomass_g_t1/6, Biomass_g),
         No_Plants = ifelse(Week == 1, 8, No_Plants)) %>%
  arrange(Cage, Week)

# infer biomass data at week 0
biomass.0.df <- merge.biomass.df %>%
  mutate(Week = Week - 1) %>%
  select(Cage, Week, Biomass_g_t1 = Biomass_g) %>%
  filter(Week == 0) %>%
  mutate(Biomass_g = Biomass_g_t1 - Biomass_g_t1/5,
         No_Plants = 8)

# merge week 0 and rest of biomass data
biomass.df <- bind_rows(filter(merge.biomass.df, Week > 0), biomass.0.df) %>%
  arrange(Cage, Week)
```

## Final adjustments

```{r}
timeseries_df <- left_join(timeseries_df, biomass.df) %>% # add plant biomass to timeseries data
  # turned logged predictors into named variables for measurement error model
  mutate(logBRBR_t0 = log(BRBR_t0),
         logBRBR_t = log(BRBR_t),
         logLYER_t0 = log(LYER_t0),
         logLYER_t = log(LYER_t),
         logLYER_t_adj = log(LYER_t_adj),
         logPtoid_t0 = log(Ptoid_t0),
         logPtoid_t = log(Ptoid_t),
         logPtoid_t_adj = log(Ptoid_t_adj)) %>%
  # add measurement error. if alive, used median observed measurement error, if not, set to small value (0.01) to allow measurement error model to extrapolate
  mutate(se_logBRBRt = ifelse(BRBR_Survival == 0 | is.na(BRBR_Survival) == T, 0.01,
                              ifelse(BRBR_Survival == 1, median(BRBR_me, na.rm = T), NA)),
         se_logLYERt = ifelse(LYER_Survival == 0 | is.na(LYER_Survival) == T, 0.01,
                              ifelse(LYER_Survival == 1, median(LYER_me, na.rm = T), NA)),
         se_logPtoidt = ifelse(Mummy_Ptoids_Survival == 0 | is.na(Mummy_Ptoids_Survival) == T, 0.01,
                               ifelse(Mummy_Ptoids_Survival == 1, median(Mummy_Ptoids_me, na.rm = T), NA)))
```


# Subset for multivariate analysis

## Aphids only

```{r}
# filter first two weeks of experiment
aphids_only_df <- filter(timeseries_df, Week %in% c(0,1))
```

## All species

I only retain data for which all species had positive abundances at the previous time step, and after all species have been added.

```{r}
# filter data where all species survived to time t
full_df <- filter(timeseries_df, BRBR_Survival == 1, LYER_Survival == 1, Mummy_Ptoids_Survival == 1) 

# tabulate variation in timeseries length for each cage
with(full_df, table(Week, Cage))
```


## LYER-Ptoid subset

```{r}
# filter data where BRBR is absent, but LYER and Ptoid survived to time t
LP_df <- filter(timeseries_df, BRBR_Survival == 0 | is.na(BRBR_Survival) == TRUE, LYER_Survival == 1, Mummy_Ptoids_Survival == 1) 

# tabulate variation in timeseries length for each cage
with(LP_df, table(Week, Cage))
```

## LYER only subset

Includes cage-time combinations where only the aphid *Lipaphis erysimi* persisted.

```{r}
# filter data so only LYER persists
L_df <- filter(timeseries_df, BRBR_Survival == 0 | is.na(BRBR_Survival) == TRUE, LYER_Survival == 1, Mummy_Ptoids_Survival == 0 | is.na(Mummy_Ptoids_Survival) == TRUE) 

# tabulate variation in timeseries length for each cage
with(L_df, table(Week, Cage))
```

## Ptoid only subset

Includes cage-time combinations where both aphids went extinct and the parasitoid *Diaeretiella rapae* survived for 1 to 2 more sampling periods.

```{r}
# filter data where only ptoid was transiently present in absence of other species
P_df <- filter(timeseries_df, BRBR_Survival == 0 | is.na(BRBR_Survival) == TRUE, LYER_Survival == 0 | is.na(LYER_Survival) == TRUE, Mummy_Ptoids_Survival == 1) 

# tabulate variation in timeseries length for each cage
with(P_df, table(Week, Cage))
```

# Final checks

```{r}
# all looks good, no NA or zero values
summary(select(timeseries_df, BRBR_t0, BRBR_t, BRBR_t1)) 
summary(select(timeseries_df, LYER_t0, LYER_t, LYER_t1, LYER_t_adj, LYER_t1_adj)) 
summary(select(timeseries_df, Ptoid_t0, Ptoid_t, Ptoid_t1, Ptoid_t_adj, Ptoid_t1_adj))
```


# Sava data for structural stability analyses

```{r}
# saved on June 9, 2021
# save(aphids_only_df, full_df, LP_df, L_df, P_df, file = "output/time-series-data.RData")
```

